```{r}
library(ggplot2)

src_files <- list.files("R", full.names = TRUE)
for (f in src_files) {
  source(f)
  cat(sprintf('Sourced file: %s\n', f))
}
```
# Simulate data set
```{r}
npercond <- c(20, 20, 20, 20)
crosstab <- matrix(npercond, 2, 2)

for (delta in seq(0.2, 1.2, 0.1)) {
  print(delta)
  simdata <- simulate_microarray(
    8000, sum(npercond), crosstab,
    delta = delta, phi = 0.1, zeta = 1.5,
    dropout = FALSE
  )

  file1 <- sprintf('../data/simulated/small/balanced/bal-%.1f.pdf', delta)
  write.table(simdata$X, file1, quote = FALSE, sep = '\t', row.names = F)

  ax_pca <- ggplot_pca(simdata$X, simdata$metadata, col = 'batch', pch = 'class')
  file2 <- sprintf('~/Dropbox/tmp/pca_bal-%.1f.pdf', delta)
  ggsave(file2, ax_pca, width = 6, height = 4)

  ax_top_pc <- ggplot_top_pc(
    simdata$X, simdata$metadata,
    x_axis = 'batch', col = 'batch', pch = 'class'
  )
  file3 <- sprintf('~/Dropbox/tmp/top_pc_bal-%.1f.pdf', delta)
  ggsave(file3, ax_top_pc, width = 6, height = 4)

  file4 <- sprintf('~/Dropbox/tmp/log_beta_bal-%.1f.pdf', delta)
  pdf(file4, width = 6, height = 4)
  hist(simdata$log_beta[, 1])
  dev.off()
}
```
# Load simulated data
```{r}
dir <- '../data/simulated/small/balanced'
files <- list.files(dir, full.names = T)
files <- files[c(9, 1:8, 10:11)]
print(files)
list_data <- lapply(files, read.table, sep = "\t", header = T)

ids <- sapply(files, function(x) substring(x, 34), USE.NAMES = F)
print(ids)
names(list_data) <- ids

# metadata
ncond <- 20
batch <- as.factor(rep(1:2, each = ncond * 2))
class <- rep(rep(LETTERS[1:2], each = ncond), 2)
metadata <- data.frame(batch, class, row.names = colnames(list_data[[1]]))

print(names(list_data))
X <- list_data[[2]]
```
# Data simulation
- Orthogonality between batch effects and class effects
  - Features with both batch effects and class effects results in non-orthogonality
  - Visualisation: Random unit vectors contained in a unit sphere
  - By chance, class effect genes may overlap with genes with high batch effects, resulting in non-orthogonality
  - TODO: Increase sample variation (PCA) without too much feature variability
    - Each sample has a multiplicative factor that is normalised
```{r}
nclass <- 5
nbatch <- 10
npercond <- 5
# n_kg <- rep(npercond, nbatch * nclass)
crosstab <- matrix(npercond, nclass, nbatch)
print(crosstab)

delta <- 2
simdata <- simulate_microarray(
  10000, crosstab,
  delta = delta, gamma = 1,
  phi = 0.1, zeta = 1.5,
  epsilon = 0.5,
  kappa = 0.2,
  dropout = FALSE
)
```
# Plot
## PCA
```{r}
ax_pca <- ggplot_pca(
  simdata$X, simdata$metadata,
  col = 'batch', pch = 'class'
)

file <- "tmp/pca-simulated_b10c5_delta20.png"
ggsave(file, ax_pca, width = 5, height = 5)
```
## Feature
```{r}
i <- 0
i <- i + 1
feature <- data.frame(
  value = as.vector(data.matrix(simdata$X[i, ])),
  batch = simdata$metadata$batch,
  class = simdata$metadata$class
)

ax <- ggplot(feature) +
  geom_point(
    aes(x = batch, y = value, pch = class, col = batch),
    position = position_jitterdodge(jitter.width = .5),
    show.legend = FALSE
  ) +
  ylim(0, 15)

file <- "tmp/feature-simulated.png"
ggsave(file, ax, width = 5, height = 5)
```
## Normalised data
```{r}
stopifnot(sum(simdata$X[simdata$X<0]) == 0)
X <- data.frame(2 ^ simdata$X)
X_scaled <- log2_transform(normaliseMeanScaling(X))

ax_pca <- ggplot_pca(
  X_scaled, simdata$metadata,
  col = 'batch', pch = 'class'
)
file <- "tmp/pca-simulated_scaled.png"
ggsave(file, ax_pca, width = 5, height = 5)

i <- 1
feature <- data.frame(
  value = as.vector(data.matrix(X_scaled[i, ])),
  batch = simdata$metadata$batch,
  class = simdata$metadata$class
)

ax <- ggplot(feature) +
  geom_point(
    aes(x = batch, y = value, pch = class, col = batch),
    position = position_jitterdodge(jitter.width = .5),
    show.legend = FALSE
  ) +
  ylim(0, 15)
file <- "tmp/feature-simulated_scaled.png"
ggsave(file, ax, width = 5, height = 5)
```
## Colsum
```{r}
l1 <- colSums(simdata$X)
feature <- data.frame(
  value = l1,
  batch = simdata$metadata$batch,
  class = simdata$metadata$class
)

ax <- ggplot(feature) +
  geom_point(
    aes(x = batch, y = value, pch = class, col = batch),
    position = position_jitterdodge(jitter.width = .5),
    show.legend = FALSE
  )
file <- "tmp/l1-simulated.png"
ggsave(file, ax, width = 5, height = 5)
```
## Histogram
```{r}
hist(simdata$Z[1, ], breaks = 20)
```